---
title: "Regresion Models Project"
author: "Abdelrahman Elsehaily"
date: "October 7, 2017"
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.width=14, fig.height=6) 
```
## 1.Executive Summary
This is the project of the Regression models course, the goal of the project is to explore the relationship between a set of variables and miles per gallon (MPG) (outcome) in the motor cars dataset, we are particularly interested in the following two questions:  
*“Is an automatic or manual transmission better for MPG”  
*"Quantify the MPG difference between automatic and manual transmissions"  
First I used t-test to see if there is significant difference between manual and automatic then used linear regression to quntify this difference then adjusted that model to enhance the **r-squared** and used **anova** to compare the different model.

## 2.Analysis

### Is the difference between maual and automatice is significant?
As seen in figure 2 automatic has  mpg lower values than manual so we want to check if that difference is significant.  
```{r}
mtcars$am<-ifelse(mtcars$am==0,"automatic","manual")
t.test(mpg~am,data = mtcars)
```
The p-value is lower than 0.0025 (tow-sided test) so there is a significant difference in the mean of the two groups.

### Quntifying the difference between the manual and automatic
using am only implement linear regression model
```{r}
amModel<-lm(mpg~ am,data=mtcars)
summary(amModel)
```
from the summary autmatic cars has 17.147 mpg **on average** and manual cars is higher with value 7.245

### Adjustment  
The adjusted R-squared is 0.3385 which mean that am could only explain 33.8 % of the variance, so we should add another variables to our model but that should depend on the correlation between the mpg and the other variables (see appendix 2)

```{r}
allCorModel<-lm(mpg~ am+wt+qsec,data=mtcars)
allVarModel<-lm(mpg~ .,data=mtcars)
anova(amModel,allCorModel,allVarModel)
```
The Residual Sum Squared (RSS) has decreased from 720 to 169.29  in the second one and as **p-value<0.05** this model is **significantly** better than the one with the variable am only.  
In the last model the Residual saured error has decreased but the p-value is bigger than 0.05 so this model is not significantly better than the prevoius so I think using all variables may have multicolinearity problem represent an overfitted model.

### Diagnostics
checking the Varience inflation factor to see how the variable affect each other in the last two model
```{r}
library(car)
vif(allCorModel)
vif(allVarModel)
```
so you see how large the vif in the last model that because of the multicolinearity issue also our chosen model has vif bigger than 2 but it is relatively low compared to the last model.


## Appendix

### Appendix 1:The difference in mpg with manual and automatice
```{r fig.height=5}
library(ggplot2)
ggplot(data = mtcars,aes(x=as.factor(am),y=mpg,fill=as.factor(am)))+geom_boxplot()
```

### Appendix 2: correlation between the mtcars columns
```{r}
data("mtcars")
library(corrplot)
corrplot(cor(mtcars[, ]))
```


### Appendix 3:The dignostic plots of the selected model  
seems from the plot that there's not strong pattern in the residuals
```{r}
par(mfrow=c(2,2))
plot(allCorModel)
```