---
title: "Reproducabl Research Final Project"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE,fig.width=6, fig.height=3) 

```

# Analyzing the effects of weather events on United states

## 1.synopsis 

This project is a par Reproducable Research course which is a part of the Data Science Specialization on Coursera, The basic goal of this assignment is to explore the [NOAA Storm Database](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2) and determining which events has the most effect on health and economic, the data is collected from 1950 and end in November 2011
the analysis doesn't take into the consideration the dates or the states but the total effects on health, property and crops caused by the different weather events on United States.


## 2. Data Processing

### Importing dpylr for wrangling the data, and ggplot2 for visualization
```{r}
library("dplyr")
library("ggplot2")
library("gridExtra")
```
#### Reading data
```{r cache=TRUE}
rawdata=read.csv("repdata%2Fdata%2FStormData.csv.bz2")
head(rawdata,3)
```
### Dealing with the eponents 
from reading the documentation I figured out that the coulumn "PROPDMGEXP", "CROPDMGEXP" is eponent for the "PROPDMG", "CROPDMG" wicht represents the damages in money loss
```{r}
levels(rawdata$PROPDMGEXP)
```
```{r}
levels(rawdata$CROPDMGEXP)

```

creating a new data frames that represnt the exponents with the real numbers and merging them to the data 
```{r cache=TRUE}
propdmgdf=data.frame(PROPDMGEXP=levels(rawdata$PROPDMGEXP),PROPDMGEXPNUM= c(0,0,0,0,1,1,10,100,1000,1e+05,1e+06,1e+07,1e+08,1e+09,100,100,1000,1e+06,1e+06))
cropdf=data.frame(CROPDMGEXP=levels(rawdata$CROPDMGEXP),CROPDMGEXPNUM= c(1,0,1,10,1e+09,1000,1000,1e+06,1e+06))
numericdf=merge(rawdata,propdmgdf,by="PROPDMGEXP")
numericdf=merge(numericdf,cropdf,by="CROPDMGEXP")
```
Creating new cloumns represents the multiblcations of the damages and the exponents
```{r}
numericdf=numericdf%>%mutate(TOTALPROPDMG=PROPDMGEXPNUM*PROPDMG,TOTALCROPDMG=CROPDMGEXPNUM*CROPDMG)
```
### getting the required data frame
first selected the required coulmns only for the results part, then grouped by event type to get the total (injuries, fatalities, crop damages, Property damages)
```{r}
analysisdata=select(numericdf,
                  EVTYPE,TOTALPROPDMG,TOTALCROPDMG,INJURIES,FATALITIES)%>%
                  group_by(EVTYPE)%>%
                  summarise(TOTALPROPDMG=sum(TOTALPROPDMG),
                            TOTALCROPDMG=sum(TOTALCROPDMG),
                            INJURIES=sum(INJURIES),
                            FATALITIES=sum(FATALITIES))
```
## 2.Results

### Health Effect
getting the first ten event types with highest deaths and highest injuries in distinct plots
```{r}
p1<-ggplot(data=slice(arrange(analysisdata,desc(FATALITIES)),1:10),aes(x=reorder(EVTYPE,-FATALITIES),y=FATALITIES))+geom_bar(stat = "identity")+theme(axis.text.x = element_text(angle = 90, hjust = 1))+xlab("")+ggtitle("Highsest Events FATALITIES")
p2<-ggplot(data=slice(arrange(analysisdata,desc(INJURIES)),1:10),aes(x=reorder(EVTYPE,-INJURIES),y=INJURIES))+geom_bar(stat = "identity")+theme(axis.text.x = element_text(angle = 90, hjust = 1))+xlab("")+ggtitle("Highsest Events INJURIES")
grid.arrange(p1, p2, ncol=2)
```

### Econmic Effect
```{r}
p3<-ggplot(data=slice(arrange(analysisdata,desc(FATALITIES)),1:10),aes(x=reorder(EVTYPE,-FATALITIES),y=FATALITIES))+geom_bar(stat = "identity")+theme(axis.text.x = element_text(angle = 90, hjust = 1))+xlab("")+ggtitle("Highsest Events FATALITIES")
p4<-ggplot(data=slice(arrange(analysisdata,desc(INJURIES)),1:10),aes(x=reorder(EVTYPE,-INJURIES),y=INJURIES))+geom_bar(stat = "identity")+theme(axis.text.x = element_text(angle = 90, hjust = 1))+xlab("")+ggtitle("Highsest Events INJURIES")
grid.arrange(p3, p4, ncol=2)
```

## Conclusion

It seemd that Tornade has the highest effect on either the health and economics, also the high death doesn't imply high injuries and vise versa.